---
title: "Camera Trap Cleaning"
format: html
execute:
  eval: false
---

## Purpose

Read camera-trap detections, standardize core columns, construct scientific
names, and keep identified records.

Source files:

- `../mia-occupancy/R/io_camera_trap.R`
- `../mia-occupancy/R/taxonomy.R`
- `../mia-occupancy/R/filtering.R`
- `../mia-occupancy/scripts/run_camera_only_example.R`

These paths assume `mia-tutorials` and `mia-occupancy` are sibling
repositories. Modify `source_repo` below if your checkout uses a different
layout.

```{r}
library(here)
library(devtools)
library(dplyr)

source_repo <- here::here("..", "mia-occupancy")
load_all(source_repo)

camera_path <- here::here("data", "mia_occupancy", "camera_trap_detections_excerpt.csv")
```

```{r}
cam_raw <- read_camera_trap_data(camera_path, tz = "UTC")
glimpse(cam_raw)
```

```{r}
cam_clean <- cam_raw |>
  build_scientific_name() |>
  filter_identified_records()

cam_clean |>
  count(scientific_name, common_name, sort = TRUE) |>
  head(15)
```

Review points:

- timestamp parsing creates `photo_datetime` and `photo_date`;
- camera name is standardized to `site_id`;
- longitude/latitude become lowercase fields;
- taxonomy remains inspectable before modeling.

