---
title: "Events, Periods, and Detection History"
format: html
execute:
  eval: false
---

## Purpose

Convert cleaned detections into independent events and replicate-level
detection histories.

Source files:

- `../mia-occupancy/R/events.R`
- `../mia-occupancy/R/effort.R`
- `../mia-occupancy/R/detection_history.R`
- `../mia-occupancy/R/utils_dates.R`

These paths assume `mia-tutorials` and `mia-occupancy` are sibling
repositories. Modify `repos_root` or `source_repo` if your checkout uses a
different layout.

```{r}
library(devtools)
library(dplyr)

root <- normalizePath("../..", winslash = "/", mustWork = TRUE)
repos_root <- normalizePath(file.path(root, ".."), winslash = "/", mustWork = TRUE)
source_repo <- file.path(repos_root, "mia-occupancy")
load_all(source_repo)

camera_path <- file.path(root, "data", "mia_occupancy", "camera_trap_detections_excerpt.csv")

cam_clean <- read_camera_trap_data(camera_path, tz = "UTC") |>
  build_scientific_name() |>
  filter_identified_records()
```

```{r}
cam_events <- define_independent_events(cam_clean, threshold_minutes = 30)
cam_periods <- add_sampling_period(cam_events, date_col = "photo_date", temporal_scale = "month")

effort <- infer_sampling_effort(cam_periods)
detection_counts <- count_detections(cam_periods)
taxonomy <- build_taxonomy_table(cam_periods)
completed_history <- complete_detection_history(detection_counts, effort)
detection_long <- build_detection_long(completed_history = completed_history, taxonomy = taxonomy)
detection_wide <- build_detection_wide(detection_long)
```

```{r}
list(
  raw_rows = nrow(cam_clean),
  independent_events = nrow(cam_events),
  site_period_effort_rows = nrow(effort),
  detection_long_rows = nrow(detection_long),
  detection_wide_rows = nrow(detection_wide)
)
```

```{r}
detection_long |>
  count(scientific_name, detection, sort = TRUE) |>
  head(20)
```

Important modeling caveat: `infer_sampling_effort()` infers effort from
site-period combinations with retained records. Replace this with deployment
metadata when true camera effort is available.

