---
title: "spOccupancy Input and Covariates"
format: html
execute:
  eval: false
---

## Purpose

Assemble the final list shape expected by `spOccupancy`: `y`, `occ.covs`,
`det.covs`, and `coords`.

Source files:

- `C:/Users/dev/work/manglaria/repos/mia-occupancy/R/spoccupancy_input.R`
- `C:/Users/dev/work/manglaria/repos/mia-occupancy/R/weather.R`
- `C:/Users/dev/work/manglaria/repos/mia-occupancy/scripts/run_data_preparation.R`

```{r}
library(devtools)
library(dplyr)
library(readr)

load_all("C:/Users/dev/work/manglaria/repos/mia-occupancy")

root <- normalizePath("../..", winslash = "/", mustWork = TRUE)
camera_path <- file.path(root, "data", "mia_occupancy", "camera_trap_detections_excerpt.csv")
weather_path <- file.path(root, "data", "mia_occupancy", "weather_daily_excerpt.csv")
```

```{r}
cam_periods <- read_camera_trap_data(camera_path, tz = "UTC") |>
  build_scientific_name() |>
  filter_identified_records() |>
  define_independent_events(threshold_minutes = 30) |>
  add_sampling_period(date_col = "photo_date", temporal_scale = "month")

effort <- infer_sampling_effort(cam_periods)
detection_counts <- count_detections(cam_periods)
taxonomy <- build_taxonomy_table(cam_periods)
completed_history <- complete_detection_history(detection_counts, effort)
detection_long <- build_detection_long(completed_history = completed_history, taxonomy = taxonomy)

sp_input <- build_spoccupancy_input(
  detection_long = detection_long,
  raw_site_data = cam_periods
)
```

```{r}
str(sp_input, max.level = 2)
dim(sp_input$y)
head(sp_input$coords)
```

## Optional weather covariates

The source `read_weather_data()` currently expects Excel input. The tutorial
fixture is CSV, so for this walkthrough we read the CSV directly and then
standardize the fields expected by `aggregate_weather_covariates()`.

```{r}
weather_long <- readr::read_csv(weather_path, show_col_types = FALSE) |>
  rename(site_id = codigo_del_sitio) |>
  mutate(
    weather_date = as.Date(weather_ts, format = "%d/%m/%Y")
  )

weather_covariates <- c("temp_median", "humidity_median", "rain_median")

weather_periods <- aggregate_weather_covariates(
  weather_data = weather_long,
  covariate_cols = weather_covariates,
  temporal_scale = "month",
  min_year = 2024
)

weather_periods |> head()
```

Use weather covariates only after confirming site IDs and sampling periods
align with camera deployments.

