---
title: "LiDAR Input Inspection"
format: html
execute:
  eval: false
---

## Purpose

Inspect LiDAR area inputs and lightweight tutorial fixtures before running any
batch point-cloud processing.

This tutorial is based on:

- `../mia-lidar/R/helpers.R`
- `../mia-lidar/R/01_inspect_inputs.R`
- `../mia-lidar/notes/003_tp_gp_ap_division_of_labor.md`
- `../mia-lidar/notes/009_dji_l1_vendor_las_quirks.md`

These paths assume `mia-tutorials` and `mia-lidar` are sibling repositories.
Modify `lidar_repo` below if your checkout uses a different layout.

Raw `.las` / `.laz` files are intentionally not copied into this tutorial repo.
The copied fixtures are a reserve polygon, an extent polygon, and a tiny
already-derived CHM GeoTIFF.

For raw point-cloud runs, download the case-study folder from Google Cloud
Storage or read it from a mounted/cloud-accessible location. Replace the
placeholder below with the current project bucket/path:

```text
gs://<manglaria-lidar-bucket>/<reserve-or-case-study>/
```

```{r}
library(here)
library(sf)
library(terra)

lidar_repo <- here::here("..", "mia-lidar")
data_dir <- here::here("data", "mia_lidar")

rbmn <- sf::st_read(here::here("data", "mia_lidar", "rbmn.geojson"), quiet = TRUE)
extent <- sf::st_read(here::here("data", "mia_lidar", "rbmnn_extent.geojson"), quiet = TRUE)
chm <- terra::rast(here::here("data", "mia_lidar", "sample_chm_mean_10m.tif"))
```

## Inspect copied fixtures

```{r}
rbmn
sf::st_bbox(rbmn)
```

```{r}
extent
sf::st_bbox(extent)
```

```{r}
chm
terra::global(chm, fun = c("min", "mean", "max"), na.rm = TRUE)
```

## Inspect one raw TP/GP/AP triplet

Use this only when raw LiDAR data are available locally or mounted from cloud
storage. If using Google Cloud Storage, first sync the relevant case-study
folder into `data/mia_lidar/raw/`, then point `area_dir` at the resulting
TP/GP/AP folder. The `data/mia_lidar/raw/` folder should stay out of git.

```{r}
source(here::here("..", "mia-lidar", "R", "helpers.R"))

lidar_root <- here::here("data", "mia_lidar", "raw", "<reserve-or-case-study>")
area_dir <- here::here("data", "mia_lidar", "raw", "<reserve-or-case-study>", "<area-folder>")
triplet <- find_lidar_triplet(area_dir)
triplet
```

Checklist before batch processing:

- `TP`, `GP`, and `AP` files are all present for the area.
- CRS and units are recorded.
- DJI L1 vendor quirks are acknowledged: pre-normalized Z, ASPRS class `0`,
  and scale-factor issues that affect some `lidR` paths.
