---
title: "LiDAR Input Inspection"
format: html
execute:
  eval: false
---

## Purpose

Inspect LiDAR area inputs and lightweight tutorial fixtures before running any
batch point-cloud processing.

This tutorial is based on:

- `C:/Users/dev/work/manglaria/repos/mia-lidar/R/helpers.R`
- `C:/Users/dev/work/manglaria/repos/mia-lidar/R/01_inspect_inputs.R`
- `C:/Users/dev/work/manglaria/repos/mia-lidar/notes/003_tp_gp_ap_division_of_labor.md`
- `C:/Users/dev/work/manglaria/repos/mia-lidar/notes/009_dji_l1_vendor_las_quirks.md`

Raw `.las` / `.laz` files are intentionally not copied into this tutorial repo.
The copied fixtures are a reserve polygon, an extent polygon, and a tiny
already-derived CHM GeoTIFF.

```{r}
library(sf)
library(terra)

root <- normalizePath("../..", winslash = "/", mustWork = TRUE)
data_dir <- file.path(root, "data", "mia_lidar")

rbmn <- sf::st_read(file.path(data_dir, "rbmn.geojson"), quiet = TRUE)
extent <- sf::st_read(file.path(data_dir, "rbmnn_extent.geojson"), quiet = TRUE)
chm <- terra::rast(file.path(data_dir, "sample_chm_mean_10m.tif"))
```

## Inspect copied fixtures

```{r}
rbmn
sf::st_bbox(rbmn)
```

```{r}
extent
sf::st_bbox(extent)
```

```{r}
chm
terra::global(chm, fun = c("min", "mean", "max"), na.rm = TRUE)
```

## Inspect one raw TP/GP/AP triplet

Use this only when raw LiDAR data are available locally.

```{r}
source("C:/Users/dev/work/manglaria/repos/mia-lidar/R/helpers.R")

area_dir <- "C:/Users/dev/work/manglaria/repos/mia-lidar/data/lidar/<area-folder>"
triplet <- find_lidar_triplet(area_dir)
triplet
```

Checklist before batch processing:

- `TP`, `GP`, and `AP` files are all present for the area.
- CRS and units are recorded.
- DJI L1 vendor quirks are acknowledged: pre-normalized Z, ASPRS class `0`,
  and scale-factor issues that affect some `lidR` paths.
