---
title: "LiDAR CHM and DEM 10 m Products"
format: html
execute:
  eval: false
---

## Purpose

Explain and inspect the canonical 10 m CHM and DEM product pattern.

Source scripts:

- `../mia-lidar/R/00_build_lax_indexes.R`
- `../mia-lidar/R/02_batch_chm_mean_10m.R`
- `../mia-lidar/R/03_batch_dem_mean_10m.R`
- `../mia-lidar/out/lidar/lidar/001_chm_mean_10m_data_product.md`
- `../mia-lidar/out/lidar/lidar/002_dem_mean_10m_data_product.md`

## Product logic

The CHM tutorial product reads AP point clouds, rasterizes canopy height at
1 m, and aggregates to 10 m with mean. The DEM counterpart reads GP point
clouds and interpolates local ground micro-topography before the same 10 m
aggregation step.

```{r}
library(here)
library(terra)

chm_path <- here::here("data", "mia_lidar", "sample_chm_mean_10m.tif")
chm <- terra::rast(chm_path)
chm
```

```{r}
terra::res(chm)
terra::crs(chm)
terra::global(chm, fun = c("min", "mean", "max"), na.rm = TRUE)
```

## Batch run pattern

Run from the `mia-lidar` source repo after editing each script's config block:

```{bash}
Rscript R/00_build_lax_indexes.R
Rscript R/02_batch_chm_mean_10m.R
Rscript R/03_batch_dem_mean_10m.R
```

Expected output shape:

- one single-band GeoTIFF per area,
- LAS-native CRS,
- 10 m grid,
- product names ending in `_chm_mean_10m.tif` or `_dem_mean_10m.tif`.

