---
title: "LiDAR Teacher Grid Consensus Height"
format: html
execute:
  eval: false
---

## Purpose

Explain the v1.1 consensus dominant-height teacher label and inspect copied
teacher manifests/audits.

Source references:

- `../mia-lidar/R/04_batch_teacher_grid_v1.R`
- `../mia-lidar/instructions/02_lidar_teacher_products.md`
- `../mia-lidar/notes/011_teacher_grid_v1_consensus_dominant_height.md`

The active teacher target is:

```text
H_dom_consensus = mean(chm_p95_local, zq95_local) when agreement_qc == 1
H_dom_consensus = NA otherwise
```

The current agreement rule is:

```text
abs(chm_p95 - zq95) <= max(1.5 m, 0.15 * mean(chm_p95, zq95))
```

```{r}
library(readr)
library(dplyr)

root <- normalizePath("../..", winslash = "/", mustWork = TRUE)
data_dir <- file.path(root, "data", "mia_lidar")

manifest <- readr::read_csv(file.path(data_dir, "teacher_manifest.csv"), show_col_types = FALSE)
agreement <- readr::read_csv(file.path(data_dir, "audit_agreement_per_parcel.csv"), show_col_types = FALSE)
```

```{r}
manifest |>
  count(agreement_rule_version, agreement_floor_m, agreement_ratio)
```

```{r}
agreement |>
  summarise(
    parcels = n(),
    total_cells = sum(n_total, na.rm = TRUE),
    both_present = sum(n_both, na.rm = TRUE),
    within = sum(n_within, na.rm = TRUE),
    outside = sum(n_outside, na.rm = TRUE),
    pct_outside_given_both = 100 * outside / both_present
  )
```

```{r}
agreement |>
  arrange(desc(pct_outside_given_both)) |>
  select(parcel_id, n_total, n_both, pct_outside_given_both, mean_h_p50, mean_offset) |>
  head(10)
```

## Batch run pattern

Run from `mia-lidar` only after confirming reserve paths:

```{bash}
Rscript R/04_batch_teacher_grid_v1.R
```

Archive these outputs for review:

- `teacher_labels/`
- `agreement/`
- `agreement_qc/`
- `agreement_reason/`
- `chm_p95/`
- `chm_p95_count/`
- `zq95/`
- secondary feature rasters

