---
title: "LiDAR Tolerance Recut and Agreement Audit"
format: html
execute:
  eval: false
---

## Purpose

Show how agreement diagnostics support tolerance-rule decisions, and how
rule-dependent outputs can be re-derived without reprocessing point clouds.

Source references:

- `../mia-lidar/R/05_audit_teacher_agreement.R`
- `../mia-lidar/R/06_apply_alternate_tolerance.R`
- `../mia-lidar/notes/012_architect_decision_agreement_rule_recut_v1_1.md`

```{r}
library(here)
library(readr)
library(dplyr)
library(ggplot2)

sensitivity <- readr::read_csv(
  here::here("data", "mia_lidar", "audit_tolerance_sensitivity.csv"),
  show_col_types = FALSE
)
sensitivity
```

```{r}
sensitivity |>
  arrange(floor_m, ratio) |>
  mutate(rule = paste0("floor=", floor_m, ", ratio=", ratio)) |>
  select(rule, n_both_present, n_outside, pct_outside_pool)
```

```{r}
ggplot(sensitivity, aes(x = ratio, y = pct_outside_pool, color = factor(floor_m))) +
  geom_line() +
  geom_point() +
  labs(
    x = "Relative tolerance ratio",
    y = "Outside tolerance (%)",
    color = "Floor (m)"
  )
```

## Re-derive outputs

`R/06_apply_alternate_tolerance.R` reads durable rule-independent rasters
(`chm_p95`, `zq95`, `chm_p95_count`) and rewrites only the rule-dependent
outputs:

- `agreement`
- `agreement_qc`
- `agreement_reason`
- `teacher_labels`

```{bash}
Rscript R/06_apply_alternate_tolerance.R
```

This is the preferred path for tolerance re-cuts because it avoids expensive
LiDAR rasterization.

